\ZpL: a p-adic precision package

### 2018-01-07 ZpL/ZpL.tex
First examples in ZpL.tex
First version of 5.1

\usepackage{amsmath,amssymb}
%\usepackage{amsthmnoproof}
\usepackage{amsthm}
\usepackage{mathrsfs}
%\let\bibsection\relax
%\usepackage{amsrefs}
%\usepackage[usenames,dvipsnames]{color}
\usepackage{stmaryrd}
\usepackage{enumerate}
%\usepackage[algoruled,vlined,english,linesnumbered]{algorithm2e}
%\usepackage[pdfpagelabels,colorlinks=true,citecolor=blue]{hyperref}
\usepackage{comment}
\usepackage{multirow}
\usepackage{xspace}
%\usepackage{tikz}

@@

\begin{itemize}
\item Compare $px$ and $x + x + \cdots +x$ ($p$ times)
\item Compare $x^p$ and $x \cdot x \cdots x$ ($p$ times)
\item $x$ and $y$ are given at different precision, compute $u = x+y$,
$v = x-y$ and $u+v$
\end{itemize}

@@

\begin{itemize}
\item How to access to the precision
\item Diffused digits
\item Historic of variables creation/deletion, etc...
\item Other features?
\end{itemize}

@@

@@

>>>>

\usepackage{amsmath,amssymb}
%\usepackage{amsthmnoproof}
\usepackage{amsthm}
\usepackage{mathrsfs}
%\let\bibsection\relax
%\usepackage{amsrefs}
%\usepackage[usenames,dvipsnames]{color}
\usepackage{stmaryrd}
\usepackage{enumerate}
%\usepackage[algoruled,vlined,english,linesnumbered]{algorithm2e}
%\usepackage[pdfpagelabels,colorlinks=true,citecolor=blue]{hyperref}
\usepackage{comment}
\usepackage{multirow}
\usepackage{xspace}
\usepackage{tabularx,multicol}
%\usepackage{tikz}

@@

\begin{tabular}{|l|l|}
\verb?R = Zp(3,5)? & \verb?R = ZpL(3,5)? \\
\verb?x = R(143,5); x? & \verb?x = R(143,5); x? \\
\hfill\verb?...12022? & \hfill\verb?...12022? \\
\verb?3*x? & \verb?3*x? \\
\hfill\verb?...120220? & \hfill\verb?...120220? \\
\verb?x + x + x? & \verb?x + x + x? \\
\hfill\verb?...20220? & \hfill\verb?...120220? \\
\verb?x^3? & \verb?x^3? \\
\hfill\verb?...020222? & \hfill\verb?...020222? \\
\verb?x * x * x? & \verb?x * x * x? \\
\hfill\verb?...20222? & \hfill\verb?...020222? \\
\end{tabular}

\begin{tabular}{|l|l|}
\verb?R = Zp(2,10)? & \verb?R = ZpL(2,10)? \\
\verb?x = R(987,10); x? & \verb?x = R(987,10); x? \\
\hfill\verb?...1111011011? & \hfill\verb?...1111011011? \\
\verb?y = R(21,5); y? & \verb?y = R(21,5); y? \\
\hfill\verb?...10101? & \hfill\verb?...10101? \\
\verb?u = x + y; u? & \verb?u = x + y; u? \\
\hfill\verb?...10000? & \hfill\verb?...10000? \\
\verb?v = x - y; v? & \verb?v = x - y; v? \\
\hfill\verb?...00110? & \hfill\verb?...00110? \\
\verb?u + v? & \verb?u + v? \\
\hfill\verb?...10110? & \hfill\verb?...11110110110? \\
\end{tabular}

@@

\begin{itemize}
\item Discuss coercion/conversion: \texttt{Zp} coerces to \ZpL but no coersion
in the other direction
\item How to access to the precision
\item Diffused digits
\item Historic of variables creation/deletion, etc...
\item Other features?
\end{itemize}

@@

\begin{tabular}{|l|l|}
\verb?R = Zp(2)? & \verb?R = ZpL(2)? \\
\multicolumn{2}{|l|}{\tt x, y, z, t = R(1,15), R(1,15), R(1,15), R(3,15)} \\
\multicolumn{2}{|l|}{\tt for \_ in range(20):} \\
\multicolumn{2}{|l|}{\tt \ \ \ \ x, y, z, t = y, z, t, (y*t + z*z)/x} \\
\multicolumn{2}{|l|}{\tt \ \ \ \ print t} \\
\hfill\verb?...000000000000100? & \hfill\verb?...000000000000100? \\
\hfill\verb?...000000000001101? & \hfill\verb?...000000000001101? \\
\hfill\verb?...000000000110111? & \hfill\verb?...000000000110111? \\
\hfill\verb?...101010111010111? & \hfill\verb?...101010111010111? \\
\hfill\verb?...1100111101111? & \hfill\verb?...101100111101111? \\
\hfill\verb?...1110000010010? & \hfill\verb?...111110000010010? \\
\hfill\verb?...0001000111001? & \hfill\verb?...100001000111001? \\
\hfill\verb?...0000011111101? & \hfill\verb?...100000011111101? \\
\hfill\verb?...1000000110101? & \hfill\verb?...001000000110101? \\
\hfill\verb?...101101010011? & \hfill\verb?...010101101010011? \\
\hfill\verb?...110000000000? & \hfill\verb?...001110000000000? \\
\hfill\verb?...000101011101? & \hfill\verb?...111000101011101? \\
\hfill\verb?...001001101011? & \hfill\verb?...111001001101011? \\
\hfill\verb?...000011110011? & \hfill\verb?...111000011110011? \\
\hfill\verb?...11? & \hfill\verb?...0000000111? \\
\hfill\verb?...10? & \hfill\verb?...1010001110? \\
\hfill\verb?...01? & \hfill\verb?...1110110001? \\
\hfill\verb?...01? & \hfill\verb?...011110111001? \\
\hfill\verb?...01? & \hfill\verb?...100101101001001? \\
\hfill\verb?...1? & \hfill\verb?...11111000011? \\
\end{tabular}

@@

\subsection{Exploration}

The purpose of this package is that of exploration
of the $p$-adic precision.
We conclude with a last example coming
from the field of the computation of
isogenies between elliptic curves over finite fields.
Techniques to compute these isogenies
through the solving of $p$-adic differential equations
have a long history, and a first
analysis of the behaviour of precision using
lattices has been investigated in \cite{LV16}.

In that article, the authors were interested in
solving $y'^2=g \times h(y)$ with
$g,h,y \in \mathbb{Z}_p \llbracket x \rrbracket$
such that $g(0)=h(0)=1$ and $y(0)=0.$

The main result was that the intrinsic loss in precision
when computing the coefficient $x^n$ of $y$
from $g$ and $h$ was in $\log_p(n)$
even though a naive analysis of
the Newton method for solving the equation
yield a loss in $\log_p(n)^2.$

However, this does not encompass the case of 
elliptic curves defined and isogeneous 
over $\mathbb{Q},$ which can yield a differential 
equation with a solution $y$ whose coefficients
have negative valuation.

For instance, the curves 
$E: \: y^2=x^3+1$
and $E': \: y^2=+ (m^2-1)x^2/4 + m^6$ are isogeneous
over $\mathbb{Q}$ if  $m$ is an odd integer.
We take $m=67.$ The corresponding isogeny has coefficients with possibly
negative $3$-adic valuation.

We can solve numerically 
the corresponding differential equation
using the algorithm of \cite{Lercier-Sirvent:08}
over $\mathbb{Q}_3$ with lattice precision.
Any generic implementation of the Newton
method of \cite{Lercier-Sirvent:08} is enough.

\begin{tabular}{|l|}
\verb?    A = QpL(3)? \\
\verb?    R.<x> = PolynomialRing(A)?\\
\verb?    m=67?\\
\verb?    g=1+(1/4)*x^2+x^6?\\
\verb?    h=1+((m^2)/4)*x^2+m^6* x^6?\\
\verb?    y=Lercier_Sirvent(g,h,64)?\\
\verb?    y[31]?\\
\verb?(467995903278061284942799/3^14 + O(3^36))*x^31?\\
\end{tabular}

We can then see that up to $63,$ the behaviour
of the loss in relative precision is the same
as for the coefficient in $x^{31}$: \textbf{no loss
occured}.
Thus, those simple manipulations
hint at a different behaviour for the loss 
in precision in at least some special cases of
of isogenies defined over $\mathbb{Q}$
but not necessarily over $\mathbb{Z}_p.$

\bibliographystyle{plain}
\bibliography{ZpL}
